Teaching Dossier

### 2019-11-17 CUT/teaching-philosophy.tex
drafting lecture section of teaching strategies

\noindent$\cdot$\emph{CO $456$: Introduction to Game Theory}

@@

In the $2017$ and $2016$ offerings of the course there was a programming project component where the students would work in groups, using their knowledge of game theory to code AI competitors for a simulated tournament. During these offerings I was responsible for programming the tournament simulation, and making sure the competitions which were held throughout the term ran smoothly.

\noindent$\cdot$\emph{CO $353$: Computational Discrete Optimization}

@@

CO $353$ covers the algorithmic aspects of discrete optimization. This includes problem formulations, greedy algorithms, local-search heuristics, approximation algorithms, linear programming duality, cutting planes, branch-and-bound, column generation, dynamic programming, problem reductions and NP-hardness.

\noindent$\cdot$\emph{CO $351$: Network Flow Theory}

@@

Network Flow Theory covers the fundamental properties and algorithms for flows on directed graphs. It covers shortest path problems, the max-flow min-cut theorem and applications, minimum cost flow problems, network simplex and primal-dual algorithms.

\noindent$\cdot$\emph{CO $250$: Introduction to Optimization}

@@

This is the second course in the calculus sequence at Waterloo for science students, i.e.\ students in physics, chemisty, biology, and economics. This course is typically the last calculus course for many science students and tries to cover a broad range of topics. It teachs the evaluation of integrals, applications to volumes and arc lengths, differential equations, sequences and series, Taylor's Theorem, parametric and vector representations of curves, and polar coordinates in the plane.

@@

\section{Teaching Strategies}
%reread philosophy statement and pick out teaching strategies to describe in detail
%write a stream of consciousness about teaching strategy that you can later organize thoughts from
%consider whether it is better to organize by teaching context or by technique

\section{Evaluation of Teaching}
I find feedback on my teaching to be tremendously valuable. It is the easiest way for me to identify potential avenues for improvement, while at the same time motivating me by highlighting strong points of my teaching style. I have organized my evaluations below into two categories: those from students in courses I have taught, and those from peers and teaching professionals observing my lectures.
\subsection{Student Evaluations}
%Chart out survey responses for CO 327 with survey adjectives for context for scores (some here, rest in appendix)
%Student comment strong points
%Student comment targets for change
\subsection{Teaching Observations}
%highlights from 1st observation
%highlights from 2nd observation with a narrative about improving

>>>>

\noindent\emph{CO $456$: Introduction to Game Theory}

@@

In the $2017$ and $2016$ offerings of the course there was a programming project component where the students would work in groups, using their knowledge of game theory to code AI competitors for a simulated tournament. During these offerings I was responsible for programming the tournament simulation, and making sure the competitions which were held throughout the term ran smoothly.
\bigskip

\noindent\emph{CO $353$: Computational Discrete Optimization}

@@

CO $353$ covers the algorithmic aspects of discrete optimization. This includes problem formulations, greedy algorithms, local-search heuristics, approximation algorithms, linear programming duality, cutting planes, branch-and-bound, column generation, dynamic programming, problem reductions and NP-hardness.
\bigskip

\noindent\emph{CO $351$: Network Flow Theory}

@@

Network Flow Theory covers the fundamental properties and algorithms for flows on directed graphs. It covers shortest path problems, the max-flow min-cut theorem and applications, minimum cost flow problems, network simplex and primal-dual algorithms.
\bigskip

\noindent\emph{CO $250$: Introduction to Optimization}

@@

This is the second course in the calculus sequence at Waterloo for science students, i.e.\ students in physics, chemisty, biology, and economics. This course is typically the last calculus course for many science students and tries to cover a broad range of topics. It teachs the evaluation of integrals, applications to volumes and arc lengths, differential equations, sequences and series, Taylor's Theorem, parametric and vector representations of curves, and polar coordinates in the plane.
\bigskip

@@

\section{Teaching Strategies}
%reread philosophy statement and pick out teaching strategies to describe in detail:
%       questioning strategies in lecture
%       outlining, learning outcomes, reviews
%       problem solving frameworks
%write a stream of consciousness about teaching strategy that you can later organize thoughts from
%consider whether it is better to organize by teaching context or by technique
\subsection{In Lecture}
\paragraph{}
The lecture is the fundamental unit of instruction in a university setting. As such I place a great emphasis on preparing my lectures in a way that enables my students to extract as much benefit as possible from each lecture. For me, the purpose of a lecture is to give my students learning outcomes, sufficient motivation to attain them, and all the tools necessary to make achieving those learning outcomes as easy as possible.

\paragraph{Learning Outcomes}Everything I do in lecture is centre around learning outcomes. For each lecture, depending on time and where we are in the course, I will identify one to four learning outcomes I want the students to get from the lecture. I will write them down on the leftmost board in the room, leave them up for the entire session, and check them off as we go. My learning outcomes are always specific, descriptive, and measurable. For example, not ``learn Braess's paradox'' but ``Braess's paradox: demonstrate how adding a road to a traffic network can increase total congestion''. Structuring the session around learing outcomes and presenting them at the start serves to prime students' brains for learning, clarify expectations for them, and provide a helpful structure to the lecture allowing students to organize their thoughts and measure progress.
\paragraph{Past, Present, and Future} Each lecture does not exist in a vacuum. Typically their is a lecture which will proceed it, and  a lecture which will follow it. When preparing a lecture, I reflect on which content from previous lectures will be needed to achieve the day's learning outcomes, and where the learning outcomes of the day will be needed in future lectures. At the start of lecture I add bullets to the outline containing the learnign outcomes describing the skills from previous lectures which will be relevant today, and bullets describing how the content of the day will feed into future goals in the course. This serves to put the material in context, which supports student retention. It allows them to form connections between what they learned last class, or last week, or maybe even last month and what they are working on today. Furthermore, the mention of future goals motivates them to study the material of this session and mentally prepares them for what is coming next.

With the past, present, and future learning outcomes fleshed out, my outline is complete. I will make reference to this outline throuhgout the lecture, as the relevant pieces come up. Further, I save time towards the end of the session to review the entire outline. Allowing me to once again summarize the learning outcomes of the day and how they fit into the bigger picture of the course.

\paragraph{Pre-test}Directly after the outline of learning outcomes, I will usually continue the lecture with a pre-test activity. The goal of a pre-test is to evaluate, with no stakes, the baseline of comprehension for the material that will be covered in the day's lecture. In cases where the lecture exists in a series on related material for which I already have a good feel for the level of understanding my students have I will skip the pre-test. That said, the pre-test is a valuable method to help me calibrate the level of my instruction to the level of the students in the room, and I only omit it as a conscious decision when I feel my teaching level is well calibrated to the students already. The pre-test does not need to take more than a couple minutes. It can be as simple as an informal show-of-hands poll on whether students have heard of the main concept of discussion today and to what extent. That is the method I favour and use most often. Another option is to give a short exercise at the start of class. This method is somewhat more time consuming. I only use it when I need to be absolutely certain the students are at a certain level of learning attainment. Say for example, I am teaching a very technical proof that day and want to be sure that students have grasped the critical lemma from the previous lecture. The pre-test is only as valuable as my willingness to adapt based on its results. This means building in some flexibility into my lecturing style. If I observe the majority of students are lacking in background, then I may have to spend more time on the background during that lecture and not get as far into the content I planned. Conversely, if the students are exceptionally prepared, I need to be willing to skip content they have previously seen and are comfortable with and be prepared to go deeper into the material than initially planned.

\paragraph{Post-test}The mirror image of the pre-test is the post-test. Partially I use the post-test to signal to myself how well the students absorbed the material from this session, but more importantly I use the post-test to signal to students how well they absorbed the material from this session. A post-test is a small exercise performed at the end of a lecture to evaluate attainment of the desired learning outcomes of that lecture. In practice, I want post-tests to be fast, taking around five minutes, and they usually will target the lower levels of Bloom's taxonomy. Usually they will feature a calculation using a technique from the day's lecture, or ask students to give another example of a concept presented today. In more proof-focused courses, they may be asked to observe a simple corollary which directly follows from a lemma presented in class. The post-tests are meant to be done individualy or in small groups. Some groups finish quickly and are free to leave early. Other groups realize they did not understand something from the lecture they initially thought they did, and it triggers questions that I can immediately address for them.

On questioning strategies and problem solving frameworks.

\subsection{Outside Lecture}

\subsection{Assessment}
\section{Evaluation of Teaching}
I find feedback on my teaching to be tremendously valuable. It is the easiest way for me to identify potential avenues for improvement, while at the same time motivating me by highlighting strong points of my teaching style. I have organized my evaluations below into two categories: those from students in courses I have taught, and those from peers and teaching professionals observing my lectures.
\subsection{Student Evaluations}
%Chart out survey responses for CO 327 with survey adjectives for context for scores (some here, rest in appendix)
%Student comment strong points
%Student comment targets for change
\subsection{Teaching Observations}
%highlights from 1st observation
%highlights from 2nd observation with a narrative about improving
